---
title: "stats"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes:
- \usepackage{amsmath}
- \usepackage{mathtools}
- \usepackage{float}
- \usepackage{xcolor,pifont}
- \newcommand{\cmark}{\Large\textcolor{green}{\ding{52}}}
- \newcommand{\xmark}{\Large\textcolor{red}{\ding{55}}}
---



Hypothesis testing test code

```{r}
require(dplyr)
require(tidyr)
require(knitr)
require(ggplot2)
require(maps)
require(RColorBrewer)
require(summarytools)
require(gridExtra)

sessionInfo()

```

```{r}

#A distinct list of beer styles. A distinct beer style is noted as having a unique style name, IBU, ABV, and Ounces values.

styles <- beer_clean %>% 
              distinct(Beer_ID, Style, IBU, ABV, Ounces) %>% 
              arrange(Style) %>% 
              na.omit(IBU, ABV)



```

```{r fig.width= 10, fig.height = 10}


lm_eqn = function(m) {

  l <- list(a = format(coef(m)[1], digits = 2),
      b = format(abs(coef(m)[2]), digits = 2),
      r2 = format(summary(m)$r.squared, digits = 3));

  if (coef(m)[2] >= 0)  {
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,l)
  } else {
    eq <- substitute(italic(y) == a - b %.% italic(x)*","~~italic(r)^2~"="~r2,l)    
  }

  as.character(as.expression(eq));                 
}

ggplot(styles, aes(x=ABV, y=IBU)) +
  geom_point(aes(colour=as.factor(Ounces))) +
  scale_colour_brewer() +
  geom_smooth(method = "lm") +
  theme(legend.position="none")
  geom_text(aes(x = .06, 
                y = 100, 
                label = lm_eqn(lm(ABV ~ IBU ,styles))), parse = TRUE, color = misc_warm)




```



# Check for Outliers in IBU
```{r fig.height = 3, fig.width = 14}



ibu_outliers <- boxplot(styles$IBU, plot = FALSE)[["out"]]

abv_outliers <- boxplot(styles$ABV, plot = FALSE)[["out"]]


x<-boxplot(styles$IBU, plot = FALSE)

bp_abv <- ggplot((styles %>% drop_na(ABV)), aes(x="", y=ABV)) +
      geom_point(aes(fill = ifelse((ABV %in% abv_outliers),"Outlier","Valid")), size = 5, shape = 21, position = position_jitter())+
      stat_boxplot(geom ='errorbar') +
      geom_boxplot(alpha=.5, outlier.shape = NA) +
      stat_summary(fun.y=mean, geom="point", shape=5, size=4) + 
      coord_flip()


bp_ibu <- ggplot((styles %>% drop_na(IBU)), aes(x="", y=IBU)) +
      geom_point(aes(fill = ifelse((IBU %in% ibu_outliers),"Outlier","Valid")),
                 size = 5, 
                 shape = 21, 
                 position = position_jitter())+
      stat_boxplot(geom ='errorbar') +
      geom_boxplot(alpha=.5, outlier.shape = NA) +
      
      coord_flip()

grid.arrange(bp_abv, bp_ibu)



```

# Check for Outliers in ABV
```{r fig.height = 3, fig.width = 14}

abv_bp_stats <- boxplot(styles$ABV, plot = FALSE)

abv_outliers <- abv_bp_stats[["out"]]
abv_bp_UIF <- abv_bp_stats[["stats"]]



ggplot((styles %>% drop_na(ABV)), aes(x="", y=ABV)) +
      #geom_point(aes(fill = ifelse((ABV>3.29*IQR(ABV)),"Outlier","Valid")), size = 5, shape = 21, position = position_jitter())+#= position_jitterdodge()) +
      geom_point(aes(fill = ifelse((ABV %in% abv_outliers),"Outlier","Valid")), size = 5, shape = 21, position = position_jitter())+
      stat_boxplot(geom ='errorbar') +
      geom_boxplot(alpha=.5, outlier.shape = NA) +#, outlier.colour = "red") +
      stat_summary(fun.y=mean, geom="point", shape=5, size=4) + 
      coord_flip()
      #geom_text(aes(label=ifelse((x>4*IQR(x)|y>4*IQR(y)),label,"")), hjust=1.1)

#boxplot(styles$ABV, plot = TRUE)



```



#check normality of full dataset and sample
```{r}

#full dataaset
ggplot(styles) +
    geom_histogram(aes(x=IBU)) +
    theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) 

ggplot(styles) +
    geom_histogram(aes(x=ABV)) +
    theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) 






# #sample
# ggplot(ss) +
#     geom_histogram(aes(x=IBU)) +
#     theme(text = element_text(size=10),
#         axis.text.x = element_text(angle=90, hjust=1)) 
# 
# ggplot(ss) +
#     geom_histogram(aes(x=ABV)) +
#     theme(text = element_text(size=10),
#         axis.text.x = element_text(angle=90, hjust=1)) 
# 
# 




```

# Check normality of log sample
```{r}


styles$IBU_log <- log(styles$IBU)
styles$ABV_log <- log(styles$ABV)

ggplot(styles) +
    geom_histogram(aes(x=IBU_log)) +
    theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) 

ggplot(styles) +
    geom_histogram(aes(x=ABV_log)) +
    #geom_density() +
    theme(text = element_text(size=10),
        axis.text.x = element_text(angle=90, hjust=1)) 




# ss$IBU_log <- log(ss$IBU)
# ss$ABV_log <- log(ss$ABV)
# 
# ggplot(ss) +
#     geom_histogram(aes(x=IBU_log)) +
#     theme(text = element_text(size=10),
#         axis.text.x = element_text(angle=90, hjust=1)) 
# 
# ggplot(ss) +
#     geom_histogram(aes(x=ABV_log)) +
#     theme(text = element_text(size=10),
#         axis.text.x = element_text(angle=90, hjust=1)) 


# log plots seem to also suggest spearman over pearson

```

# QQ Plots of full and sample datasets


```{r fig.width= 10, fig.height=10}




  #calulate line fit
  y <- quantile((styles$IBU %>% na.omit()), c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  y_int <- y[1] - slope * x[1]

  ggplot(styles, aes(sample = styles$IBU)) + 
    geom_qq(shape = 16, size = 2, alpha = 0.5) +
    geom_abline(slope = slope, intercept = y_int, colour ='red', size = 1) +
    ggtitle("QQplot of IBU") +
    theme_minimal()

  
  #calulate line fit
  y <- quantile((styles$ABV %>% na.omit()), c(0.25, 0.75))
  x <- qnorm(c(0.25, 0.75))
  slope <- diff(y)/diff(x)
  y_int <- y[1] - slope * x[1]  
  
  ggplot(styles, aes(sample = styles$ABV)) + 
    geom_qq(shape = 16, size = 2, alpha = 0.5) +
    geom_abline(slope = slope, intercept = y_int, colour ='red', size = 1) +
    ggtitle("QQplot of ABV") +
    theme_minimal()  
  


```


# Analysis using Spearman Rank-Order Correlation

    + More info on Spearman test: https://statistics.laerd.com/statistical-guides/spearmans-rank-order-correlation-statistical-guide.php

+ H~o~: $\rho= 0$
+ H~A~: $\rho\neq 0$

```{r}



a=1-.05/2


m = "spearman" # ABV and IBU are both ordinal

#plot(styles$IBU, styles$ABV)
results <- cor.test(styles$IBU, styles$ABV, method = m, conf.level = a)

results

r_sq <- results[["estimate"]][["rho"]]^2
r_sq

#qt(a, results[["parameter"]][["df"]])



results <- cor.test(ss$IBU, ss$ABV, method = m, conf.level = a)

results


r_sq <- results[["estimate"]][["rho"]]^2
r_sq

#qt(a, results[["parameter"]][["df"]])

#plot(select(ss, IBU, ABV))

#print("There is strong evidence that the ABV and IBU are correlated (p-value < 0.001).  The IBU rating accounts for 40.7% of the variation in the ABV.  While IBU and ABV certainly have a correlation, that correlation is not very strong.") #TODO: calculate conf int

#TODO: check number of ties
```

There is strong evidence that the ABV and IBU are correlated (p-value < 0.001).  At a 95% confidence level, the IBU rating accounts for 31.5% of the variation in the ABV.  While IBU and ABV certainly have a correlation, that correlation is not very strong.  We can conclude that IBU rating and ABV are associated, but only that there is an association.  No causality or extrapolation can be applied to these conclusions."

Alternative questions to ask:
+ Does the mean/median ABV of a brewery 