---
title: "Matthew Trevathan Case Study 1"
author: "Matthew Trevathan"
date: "June 25, 2018"
output: html_document
---
##Introduction
This file is for first case study for Doing Data Science MSDS 6306

##Project Directory
GitHub cloud directory: https://github.com/la-mar/DDS_Case_Study_1.git.

##Libraries Used
magrittr, dplyr, plyr

##Data Read-in
```{r, ECHO=TRUE}
beers <- read.csv("Beers.csv", header = TRUE, as.is=TRUE)
brewery <- read.csv("Breweries.csv", header = TRUE, as.is=TRUE)
```

```{r, echo=FALSE}
library(magrittr)
library(dplyr)
library(plyr)
```

#Question 1. How many breweries are present in each state?
```{r, ECHO=TRUE}

by_state <- count(brewery, "State")
```

#Question 2. Merge beer data with breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r, ECHO=TRUE}
brewery_beers <- merge(beers, brewery, by.x = "Brewery_id", by.y = "Brew_ID")
colnames(brewery_beers) <- c("Brewery ID", "Beer Name", "Beer ID", "ABV", "IBU", "Style", "Ounces", "Brewery Name", "City", "State")
head(brewery_beers)
tail(brewery_beers)
```

#Question 3. Report the number of NA's in each column.
```{r, ECHO=TRUE}
na_count <- colSums(is.na(brewery_beers))
na_count
```

#Question 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
barplot(medians_IBU$median_ibu, horiz = TRUE, cex.names=0.2, names.arg = medians_IBU$State, xlim = c(0, 70))

barplot(medians_ABV$median_abv, horiz = TRUE, cex.names=0.2, names.arg = medians_ABV$State, xlim = c(0, .075))
```{r, ECHO=TRUE}
medians_ABV <- brewery_beers %>% group_by(State) %>% summarise(median_abv = median(ABV, na.rm = TRUE)) 
medians_IBU <- brewery_beers %>% group_by(State) %>% summarise(median_ibu = median(IBU, na.rm = TRUE))
```

#Question 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r, ECHO=TRUE}
ABV_state <- brewery_beers %>% group_by(State) %>% summarise(max(ABV, na.rm = TRUE))
ABV_state[which.max(ABV_state$`max(ABV, na.rm = TRUE)`),]

IBU_state <- brewery_beers %>% group_by(State) %>% summarise(max(IBU, na.rm = TRUE))
IBU_state[which.max(IBU_state$`max(IBU, na.rm = TRUE)`),]
```

#Question 6. Summary statistics for the ABV variable.
```{r, ECHO=TRUE}
ABV_summary <- summary(brewery_beers$ABV)
ABV_summary
```

#Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
plot(brewery_beers$IBU, brewery_beers$ABV)