---
title: "MSDS 6306 Beers and Breweries Case Study"
author: "Peter Flaming"
date: "6/19/2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
## Introduction
This R markdown document gives the code and output for Case Study 1: Questions 1-7 completed for the Master of Science in Data Science graduate course Doing Data Science MSDS 6306 provided by Southern Methodist University during the Summer 2018 semester. The "beers.csv" and "breweries.csv" data used in this Case Study 1 was sourced from 558 US breweries and their listed 2410 US craft beers for sale. 

## Project Directory
These data reside in the GitHub cloud directory: https://github.com/la-mar/DDS_Case_Study_1.git. 

## Libraries Used
library(tidyverse)
library(dplyr)
library(ggplot2)

## Reading Data into R
beers <- read.csv("../DDS_Case_Study_1/data/Beers.csv", header = TRUE)
breweries <- read.csv("../DDS_Case_Study_1/data/Breweries.csv", header = TRUE)

### Question 1. How many breweries are present in each state?
#### Print the count of breweries by state from "breweries" data.frame.
```{r code for breweries summary, echo=TRUE, error=TRUE}
summary(breweries$State)
```

### Question 2. Merge beer data with breweries data. Print the first 6 observations and the last six observations to check the merged file.
#### Merge "beers" data.frame with "breweries" data.frame by the common "Brewery_ID" variable into a new data.frame "beers_breweries".
```{r code for merge, echo=TRUE, error=TRUE}
beers_breweries <- merge(beers, breweries, by.x = "Brewery_id", by.y = "Brew_ID")
```

#### Rename the variables of new "beers_breweries" data.frame.
```{r code for renaming columns, echo=TRUE, error=TRUE}
colnames(beers_breweries) <- c("Brewery_ID", "Beer_Name", "Beer_ID", "ABV", "IBU", "Stye", "Ounces", "Brewery", "City", "State")
```

#### Print first 6 observations of "beers_breweries" data.frame.
```{r code for head, echo=TRUE, error=TRUE}
head(beers_breweries)
```

#### Print last 6 observations of "beers_breweries" data.frame.
```{r code for tail, echo=TRUE, error=TRUE}
tail(beers_breweries)
```

### Question 3. Report the number of NA's in each column.
#### Print the count of NA's by column from "beers_breweries" data.frame.
```{r code for NA count, echo=TRUE, error=TRUE}
beers_breweries %>% group_by(.) %>% summarise_all(funs(sum(is.na(.))))
```

### Qustion 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
#### Compute the median ABV by State for "beers_breweries" data.frame.
```{r code for median ABV, echo=TRUE, error=TRUE}
median_abv_by_state <- beers_breweries %>% group_by(State) %>% summarise(median_abv = median(ABV, na.rm = TRUE))
```

#### Compute the median IBU by State for "beers_breweries" data.frame.
```{r code for median IBU, echo=TRUE, error=TRUE}
median_ibu_by_state <- beers_breweries %>% group_by(State) %>% summarise(median_ibu = median(IBU, na.rm = TRUE))
```

#### Plot of Median ABV for "beers_breweries" data.frame.
```{r code for ABV Bar Plot, echo=TRUE, error=TRUE}
ggplot(data = median_abv_by_state) + geom_bar(mapping = aes(x = State, y = median_abv), stat = "identity")
```

#### Plot of Median IBU for "beers_breweries" data.frame.
```{r code for IBU Bar Plot, echo=TRUE, error=TRUE}
ggplot(data = median_ibu_by_state) + geom_bar(mapping = aes(x = State, y = median_ibu), stat = "identity")
```

### Question 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
#### Print the state with the maximum ABV for "beers_breweries" data.frame.
```{r code for max ABV, echo=TRUE, error=TRUE}
max_abv <- beers_breweries %>% group_by(State) %>% arrange(desc(ABV))
```

#### Print the state with the max IBU beer for "beers_breweries" data.frame.
```{r code for max IBU, echo=TRUE, error=TRUE}
max_ibu <- beers_breweries %>% group_by(State) %>% arrange(desc(IBU))
```

### Question 6. Summary statistics for the ABV variable.
#### Compute the summary statistics of the ABV variable for the "beers_breweries" data.frame.
```{r code for ABV summary, echo=TRUE, error=TRUE}
summary(beers_breweries$ABV)
```

### Question 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
#### Scatter Plot of ABV vs IBU variables for the "beers_breweries" data.frame.
```{r code for Scatter Plot, echo=TRUE, error=TRUE}
ggplot(data = beers_breweries, mapping = aes(x = ABV, y = IBU)) + geom_point(mapping = aes(x = ABV, y = IBU)) + geom_smooth(mapping = aes(x = ABV, y = IBU))
```